Thesis Draft

### 2017-09-28 thesis/draft.Rmd
fixed for LaTeX

---
title: "Thesis Draft"
author: "Gavin Chin"
date: "21 September 2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

@@

The first method used to impute missing values is to fit generalised linear models (GLMs) at each sensor location. Estimating a GLM with a Quasi-Poisson error distribution, where specify the model as:
$$E \left[\text{Hourly_Counts} \vert \text{Sensor}, \text{HDay}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{HDay}, \text{Time}}) $$
$$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
`Time` is the time of the day, while `HDay` is the day of the week, with an additional factor level for public holidays. 
Both these variables are categorical/factors. We cannot treat time of the day as an integer or numeric because it does not have a linear relationship with counts.s

@@

For sensors with a small proportion of missing values, we use a GLM quasi-poisson model again. To improve on the previous model used, we use the model:
$$E \left[\text{Hourly_Counts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$

@@

Using the two geographically closest sensors with complete cases, we use another GLM specified as:
$$E \left[ \text{Hourly_Counts}_{\text{sensor}} \vert \text{Time}, \text{Hourly_Counts}_{\text{closest sensors}} \right] = \mu_{\text{Time}, \text{Hourly_Counts}_{\text{closest sensors}}}$$
$$\text{where } \quad  \mu_{\text{Time}, \text{Hourly_Counts}_{\text{closest sensors}}} \sim \text{Time} \times \text{Hourly_Counts}_{\text{closest sensor 1}} + \text{Time} \times \text{Hourly_Counts}_{\text{closest sensor 2}}$$

@@

```{r}
ggplot(data = NULL, aes(x = original_df %>% 
             filter(Date_Time < "2017-01-01") %>%
             .[, "Southern Cross Station"],
           y = fitted_base$`Southern Cross Station`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = brewer.pal(3, 'Dark2')[1],
                       alpha = 0.7, size = 1.5,) +
           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
                       colour = brewer.pal(3, 'Dark2')[2]) +
           theme_minimal() + theme(aspect.ratio = 1) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Actual against Predicted Counts at Southern Cross Station")

@@

However, if we try to use this same model specification at sensor locations with large proportions of missing values, the fit is poor. We use the sensor at Australia on Collins to 
```{r}
options(na.action = na.omit)
ggplot(data = NULL, aes(x = original_df[-(is.na(original_df$`Australia on Collins`)), ] %>% 
             filter(Date_Time < "2017-01-01") %>%
             .[, "Australia on Collins"],
           y = fitted_base$`Australia on Collins`[-(is.na(original_df$`Australia on Collins`))])) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = brewer.pal(3, 'Dark2')[1],
                       alpha = 0.7, size = 1.5,) +
           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
                       colour = brewer.pal(3, 'Dark2')[2]) +
           theme_minimal() + theme(aspect.ratio = 1) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Actual against Predicted Counts at Australia on Collins")

```
We can see that the model is heavily biased by the false 0 values, causing the predicted values to be underestimate (slope of the fitted line is $<1$).

>>>>

---
title: "Thesis Draft"
author: "Gavin Chin"
date: "21 September 2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

@@

The first method used to impute missing values is to fit generalised linear models (GLMs) at each sensor location. Estimating a GLM with a Quasi-Poisson error distribution, where specify the model as:
$$E \left[\text{HourlyCounts} \vert \text{Sensor}, \text{HDay}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{HDay}, \text{Time}}) $$
$$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
`Time` is the time of the day, while `HDay` is the day of the week, with an additional factor level for public holidays. 
Both these variables are categorical/factors. We cannot treat time of the day as an integer or numeric because it does not have a linear relationship with counts.s

@@

For sensors with a small proportion of missing values, we use a GLM quasi-poisson model again. To improve on the previous model used, we use the model:
$$E \left[\text{HourlyCounts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$

@@

Using the two geographically closest sensors with complete cases, we use another GLM specified as:
$$E \left[ \text{HourlyCounts}_{\text{sensor}} \vert \text{Time}, \text{HourlyCounts}_{\text{closest sensors}} \right] = \mu_{\text{Time}, \text{HourlyCounts}_{\text{closest sensors}}}$$
$$\text{where } \quad  \mu_{\text{Time}, \text{HourlyCounts}_{\text{closest sensors}}} \sim \text{Time} \times \text{HourlyCounts}_{\text{closest sensor 1}} + \text{Time} \times \text{HourlyCounts}_{\text{closest sensor 2}}$$

@@

```{r}
ggplot(data = NULL, aes(x = original_df %>% 
             filter(Date_Time < "2017-01-01") %>%
             .[, "Southern Cross Station"],
           y = fitted_base$`Southern Cross Station`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = brewer.pal(3, 'Dark2')[1],
                       alpha = 0.7, size = 1.5,) +
           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
                       colour = brewer.pal(3, 'Dark2')[2]) +
           theme_minimal() + theme(aspect.ratio = 1) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Actual against Predicted Counts at Southern Cross Station")

@@

However, if we try to use this same model specification at sensor locations with large proportions of missing values, the fit is poor. We use the sensor at Australia on Collins, where the proportion of missing values is `r round(na_prop["Australia on Collins"]*100, 2)`%. 
```{r}
options(na.action = na.omit)
ggplot(data = NULL, aes(x = original_df[-(is.na(original_df$`Australia on Collins`)), ] %>% 
             filter(Date_Time < "2017-01-01") %>%
             .[, "Australia on Collins"],
           y = fitted_base$`Australia on Collins`[-(is.na(original_df$`Australia on Collins`))])) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = brewer.pal(3, 'Dark2')[1],
                       alpha = 0.7, size = 1.5,) +
           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
                       colour = brewer.pal(3, 'Dark2')[2]) +
           theme_minimal() + theme(aspect.ratio = 1) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Actual against Predicted Counts at Australia on Collins")

```
We can see that the model is heavily biased by the false 0 values, causing the predicted values to be underestimate (slope of the fitted line is $<1$). This demonstrates the importance of ensuring outliers are removed from the data.
